---
output: github_document
---


This document shows how we calculated which flow disruptions were most commonly found in disruption cascades.

All patient and hospital information has been removed from the data. To further deidentify the patients, each case was set to start at noon (720 minutes into the day), and all disruption notes were removed.


```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Load Data

```{r, message=FALSE}
rm(list = ls())
library(knitr)
library(readtv)
library(tidyverse)

f_name = 'data/events_phi_free.csv'
```


```{r}
events = read_csv(f_name) %>% 
  group_by(Case) %>% 
  group_modify(~ mutate(.x, RelativeTime = Time - min(Time, na.rm = TRUE)))

glimpse(events)
```


## Load CPA Markers

See `plot_gen2.R` for how the markers were calculated. The source code to derive the markers was output from read-tv.
```{r, echo=F}
cpa_markers = read_rds('events-cpa-markers.RDS')
glimpse(cpa_markers)
```


```{r}
getEvent = function(case, cpt, pre = TRUE) {
  if(pre) {
    rank_fn = cume_dist
    `%cmp%` = `<`
  } else {
    `%cmp%` = `>=`
    rank_fn = min_rank
  }
  
  events %>% 
    filter(Case == case) %>% 
    filter(RelativeTime %cmp% cpt) %>% 
    filter(rank_fn(RelativeTime) == 1) %>% 
    filter(row_number(RelativeTime) == max(row_number(RelativeTime)))
}
```

```{r}
cpa_markers_events = cpa_markers %>% 
  group_by(Case, cpts) %>% 
  group_modify(~ data.frame(vals = .x$vals, 
                pre = getEvent(.x$Case, .x$cpts),
                post = getEvent(.x$Case, .x$cpts, FALSE)),
               keep = TRUE)
```


## Jump Rates

```{r}
cpa_markers_events = cpa_markers_events %>% 
  group_by(Case) %>% 
  group_modify(~ .x %>% mutate(jump_rate = vals - lag(vals, default = 0),
                               is_cascade = vals >= 3,
                               jump_type =  if_else(jump_rate > 0, 
                                                    'increases', 'decreases'),
                               start = lag(cpts, default = 0),
                               end = cpts
                               )
               )
```

```{r}
cpa_markers_events %>% 
  select(Case, cpts, vals, jump_rate, jump_type, is_cascade, start, end) %>% 
  head %>% 
  kable
```


```{r}
cpa_markers_events %>% glimpse
```

## Cascade Marker Events

```{r}
cascade_start_events = cpa_markers_events %>% 
  filter(jump_type == "increases" & is_cascade)

nrow(cpa_markers_events)
```


What percent of surgeries?
```{r}
n_distinct(cascade_start_events$Case)/
  n_distinct(cpa_markers_events$Case)
```

$$
P(pre|ET)
$$


```{r}
cascade_start_events %>% 
  ungroup %>% 
  count(pre.Event.Type) %>% 
  arrange(desc(n)) %>% 
  mutate(`ratio (%)` = n/sum(events$Event.Type == pre.Event.Type) * 100) %>% 
  head %>% 
  kable
```


$$
P(post|ET)
$$

```{r}
cascade_start_events %>% 
  ungroup %>% 
  filter(jump_type == "increases") %>% 
  count(post.Event.Type) %>% 
  arrange(desc(n)) %>% 
  mutate(`ratio (%)` = n/sum(events$Event.Type == post.Event.Type) * 100) %>% 
  head %>% 
  kable
```

$P(ET_{post}|ET_{pre})$

```{r}
cascade_start_events %>% 
  ungroup %>% 
  count(pre.Event.Type, post.Event.Type) %>% 
  arrange(desc(n)) %>% 
  head %>% 
  kable
```


$P(trigger|ET)$

```{r}
getTrigger = function(pre.RelativeTime, pre.Event.Type,
                      post.RelativeTime, post.Event.Type,
                      cpts)
  if_else(abs(pre.RelativeTime - cpts) <= 3,
          pre.Event.Type, post.Event.Type)

cascade_start_events %>% 
  mutate(Event.Type = getTrigger(pre.RelativeTime, pre.Event.Type,
                                 post.RelativeTime, post.Event.Type,
                                 cpts)) %>% 
  group_by(Event.Type) %>% 
  summarise(n = n()) %>% 
  arrange(desc(n)) %>% 
  mutate(`ratio (%)` = n/sum(events$Event.Type == Event.Type) * 100) %>% 
  head %>% 
  kable
```

Common cascade event types.

```{r}
getMarker = function(case, rel_time) 
  cpa_markers_events %>% 
  filter(Case == case & start <= rel_time & end > rel_time) %>% 
  select(c_start = start, c_end = end, is_cascade, vals) %>% 
  top_n(1)

```

```{r}
cascadeEvents = function(case) {
  cme = cpa_markers_events %>% filter(Case == case)
  
  rtSafe = function(case) {
    t_end = cme %>% 
      filter(Case == case) %>% 
      {max(.$end)}
    rt = events %>% filter(Case == case) %>% pull(RelativeTime)
    if_else(rt >= t_end, t_end - 1, rt)
  }
  
  rts = rtSafe(case)
  rts %>% 
    purrr::map(function(rt){
      cme$start %>% 
        purrr::imap(~ rt >= .x & rt <= cme$end[.y] & cme$is_cascade[.y]) %>% 
        purrr::reduce(~ .x | .y)
    }) %>% 
    unlist
}
```

```{r}
getMarker(1, 5)
```

```{r}
events_and_cascades = events %>% 
  group_by(Case) %>% 
  group_modify(~ .x %>% 
                 mutate(is_cascade = cascadeEvents(unique(.x$Case))) %>% 
                 select(-"Case"), 
               keep = TRUE) %>% 
  ungroup
```


# Which Event Types are Most Likely in Cascades

```{r}
event_counts = events %>% 
  ungroup %>% 
  count(Event.Type) %>% 
  rename(n = n)
```



```{r}
table1 = events_and_cascades %>% 
  filter(is_cascade) %>% 
  count(Event.Type) %>%
  rename(n_cascade = n) %>% 
  merge(event_counts) %>% 
  mutate(p_cascade_given_et = n_cascade/n,
         p_et = n/sum(n),
         p_et_given_cascade = n_cascade/sum(n_cascade)) %>%    
  arrange(desc(p_cascade_given_et)) 
```

```{r}
table1 %>% kable
```

```{r}
events_and_cascades %>% 
  count(is_cascade) %>%
  mutate(p = n/sum(n)) %>% 
  kable
```
